Forecasting Competition 
Spring 2026
Author: Bella Huang & Yeeun Kim

GitHub repository: https://github.com/BellaHuang03/KimHuang_ENV797_TSA_ForecastCompetition_S26

1. Set up
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(forecast)
library(smooth)
library(ggplot2)
```

2. Import datasets
```{r}
#Load Data
temp_raw <- read_excel("temperature.xlsx")
humid_raw   <- read_excel("relative_humidity.xlsx")
load_raw <- read_excel("load.xlsx")

```

3. Data Wrangling
```{r}
# Create a hourly data
# Temperature
temp_hourly <- temp_raw %>%
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  mutate(temp_avg = rowMeans(select(., starts_with("t_ws")), na.rm = TRUE)) %>%
  select(date, hr, temp_avg)
 
head(temp_hourly)

# Aggregate hourly temp to daily mean
temp_daily <- temp_hourly %>%
  mutate(date = as.Date(date)) %>%        
  group_by(date) %>%
  summarise(temp = mean(temp_avg, na.rm = TRUE))

# Humidity
humid_hourly <- humid_raw %>%
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  mutate(humid_avg = rowMeans(select(., starts_with("rh_ws")), na.rm = TRUE)) %>%
  select(date, hr, humid_avg)
 
head(humid_hourly)

# Aggregate hourly humidity to daily mean
humid_daily <- humid_hourly %>%
  mutate(date = as.Date(date)) %>%        
  group_by(date) %>%
  summarise(humidity = mean(humid_avg, na.rm = TRUE))

# Load
load_hourly <- load_raw %>%
  filter(!is.na(meter_id)) %>%          
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  select(-meter_id) %>%
  pivot_longer(
    cols      = starts_with("h"),
    names_to  = "hr_str",
    values_to = "load_kwh"
  ) %>%
  mutate(hr = as.integer(sub("h", "", hr_str))) %>%  # "h1" → 1
  select(date, hr, load_kwh) %>%
  arrange(date, hr)
 
head(load_hourly)

# Create an hourly data
hourly_data <- load_hourly %>%
  left_join(temp_hourly, by = c("date", "hr")) %>%
  left_join(humid_hourly,   by = c("date", "hr")) %>%
  arrange(date, hr)

# Create a daily data
daily_data <- hourly_data %>%
  group_by(date) %>%
  summarise(
    load_daily = mean(load_kwh,  na.rm = TRUE),
    temp_avg   = mean(temp_avg, na.rm = TRUE),
    humid_avg  = mean(humid_avg,   na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  arrange(date)

```
4. Initial plot - filtering for Jan 2005 to June 2010
```{r}
test_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))

validation_data <- daily_data %>%
  filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

ts_test1 <- msts(test_data$load_daily,
                    seasonal.periods = c(7, 365.25),
                    start = c(2005, 1))
TS_Plot <- 
  ggplot(test_data, aes(x = date, y = load_daily)) +
      geom_line() +
      labs(title = "Daily Electricity Load - Training Data",
           x     = "Date",
           y     = "Load (kWh)") +
      theme_bw()
plot(TS_Plot)

# STL decomposition (more robust than decompose() for msts)
stl_fit <- mstl(ts_test1)
 
cat("\n=== STL Decomposition ===\n")
plot(stl_fit, main = "STL Decomposition — Trend + Weekly + Annual + Remainder")

#ACF and PACF plots
par(mfrow=c(1,2))
ACF_Plot <- Acf(test_data$load_daily, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(test_data$load_daily, lag = 40)
par(mfrow=c(1,1))
```

5. Decompose, forecast with seasonal Naive model and averaging
```{r}
decompose_load <- mstl(ts_test1)
plot(decompose_load, main = "STL Decomposition - Daily Load")

# remainder
deseasonal_load <- seasadj(decompose_load)

# Plot deseasonal series
TS_deseas_plot <- 
  ggplot() +
  geom_line(aes(x = test_data$date, y = as.numeric(deseasonal_load))) +
  labs(title = "Deseasonal Daily Load",
       x     = "Date",
       y     = "Load (kWh)") +
  theme_bw()
plot(TS_deseas_plot)
```
```{r}
# Model 1: Arithmetic Mean
MEAN_seas <- meanf(y = ts_test1, h = 31)
checkresiduals(MEAN_seas)

# Model 2: Seasonal Naive
SNAIVE_seas <- snaive(ts_test1, h = 31)
checkresiduals(SNAIVE_seas)

# Fitting Models to the Deseasonal Series

# Model 3: Naive on Deseasonal Data
NAIVE_deseas <- naive(deseasonal_load, h = 31)
checkresiduals(NAIVE_deseas)

# Model 4: Simple Exponential Smoothing on Deseasonal Data
SES_deseas <- ses(deseasonal_load, h = 31)
checkresiduals(SES_deseas)

```
Comparison with validation data
```{r}
actual_val <- validation_data$load_daily
rmse <- function(actual, predicted) sqrt(mean((actual - predicted)^2))

rmse_results <- c(
  MEAN        = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE      = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas  = rmse(actual_val, as.numeric(SES_deseas$mean))
)

print(rmse_results)

```

```{r}
## Retrain SES_deseas on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start = c(2005, 1))

# Decompose full training series
decompose_full <- mstl(ts_full_training)
deseasonal_full <- seasadj(decompose_full)

# Re-fit SES on deseasonal full training data
SES_final <- ses(deseasonal_full, h = 31)
checkresiduals(SES_final)

# Final forecast for July 1–31, 2011
final_forecast <- data.frame(
  date     = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast = as.numeric(SES_final$mean)
)

print(final_forecast)
```
```{r}
## Add Seasonal Component Back to Final Forecast
# Step 1: Extract seasonal components from decomposition
seasonal_weekly <- decompose_full[, "Seasonal7"]
seasonal_annual <- decompose_full[, "Seasonal365.25"]

# Step 2: Find which days in the training data correspond to July
# so we can match the same seasonal pattern for July 2011
full_dates <- seq(as.Date("2005-01-01"), as.Date("2011-06-30"), by = "day")

july_idx <- which(format(full_dates, "%m") == "07" & 
                  format(full_dates, "%d") %in% sprintf("%02d", 1:31))

# Get the LAST July in training data (July 2010) seasonal indices
last_july_idx <- tail(july_idx, 31)

# Step 3: Extract seasonal values for those 31 days
seas_weekly_july <- as.numeric(seasonal_weekly)[last_july_idx]
seas_annual_july <- as.numeric(seasonal_annual)[last_july_idx]

# Step 4: Add back to flat SES forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SES_final$mean),
  seas_weekly    = seas_weekly_july,
  seas_annual    = seas_annual_july,
  forecast       = as.numeric(SES_final$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 5: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SES (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()

# Export for submission
write.csv(final_forecast %>% select(date, forecast),
          "Kim_Huang_FirstSubmission.csv", row.names = FALSE)

```
More models to try: SARIMA & SARIMAX
```{r}
# Join exogenous variables back to test and validation dataset
test_data <- test_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

validation_data <- validation_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

# Build xreg matrix
xreg_train <- cbind(
  temp     = test_data$temp,
  humidity = test_data$humidity
)

xreg_val <- cbind(
  temp     = validation_data$temp,
  humidity = validation_data$humidity
)
```

```{r}
# Model 5: SARIMA on Deseasonal Series
SARIMA_model <- auto.arima(
  deseasonal_load,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_model)

SARIMA_fc <- forecast(SARIMA_model, h = 31)

# Model 6: SARIMAX on Deseasonal Series
SARIMAX_model <- auto.arima(
  deseasonal_load,
  xreg          = xreg_train,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMAX_model)

SARIMAX_fc <- forecast(SARIMAX_model, h = 31, xreg = xreg_val)
```
Model comparison
```{r}
# Updated RMSE Comparison
rmse_results <- c(
  MEAN         = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE       = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas   = rmse(actual_val, as.numeric(SES_deseas$mean)),
  SARIMA       = rmse(actual_val, as.numeric(SARIMA_fc$mean)),
  SARIMAX      = rmse(actual_val, as.numeric(SARIMAX_fc$mean))
)

print(sort(rmse_results))
```

Apparently, SARIMA model has the lowest RMSE, so we will forecast with SARIMA model.

```{r}
## Retrain SARIMA on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start            = c(2005, 1))

# Decompose full training series
decompose_full    <- mstl(ts_full_training)
deseasonal_full   <- seasadj(decompose_full)

# Re-fit SARIMA on deseasonal full training data
SARIMA_final <- auto.arima(
  deseasonal_full,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_final)

# Final deseasonal forecast for July 1-31, 2011
SARIMA_final_fc <- forecast(SARIMA_final, h = 31)
```

```{r}
# Step 5: Add seasonal components back to SARIMA forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SARIMA_final_fc$mean),
  seas_weekly     = seas_weekly_july,
  seas_annual     = seas_annual_july,
  forecast        = as.numeric(SARIMA_final_fc$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 6: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SARIMA (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x     = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()

# Export for submission
write.csv(final_forecast %>% select(date, forecast),
          "Kim_Huang_SecondSubmission.csv", row.names = FALSE)
```

AI statement: AI tool was used to help us change the format of date in our data wrangling process.


