---
title: "Forecasting Competition, Spring 2026"
author: "Bella Huang & Yeeun Kim"
date: "`r Sys.Date()`"
output: pdf_document
---

# 0. Introduction

This report aims to forecast a residential customer's daily electricity demand for July 2011 as a class competition. We compared multiple time series models, and evaluated their performance using RMSE and MAPE. Hourly load data from January 2005 to June 2011 are used.

GitHub repository: <https://github.com/BellaHuang03/KimHuang_ENV797_TSA_ForecastCompetition_S26>

# 1. Set up
```{r,include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

```{r,include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(forecast)
library(smooth)
library(ggplot2)
```

## Import datasets

```{r load-data, include=FALSE}
#Load Data
temp_raw <- read_excel("temperature.xlsx")
humid_raw   <- read_excel("relative_humidity.xlsx")
load_raw <- read_excel("load.xlsx")
```

```{r preview-data}
head(temp_raw, 10)
head(humid_raw, 10)
head(load_raw, 10)
```

## Data Wrangling
The raw temperature, humidity, and load datasets were systematically processed to produce consistent hourly and daily time series for analysis. Timestamps were standardized to POSIXct format, and temperature and humidity readings from multiple weather stations were averaged at the hourly level to obtain representative measures. These hourly series were subsequently aggregated to daily averages. The load dataset was reshaped from wide to long format by transforming hourly columns into a single hour variable, ensuring alignment with the temporal structure of the weather data. The resulting hourly datasets were then merged on common date and hour fields to create a unified dataset. Finally, the merged data were aggregated to the daily level, yielding average daily load, temperature, and humidity variables for subsequent modeling and analysis.

```{r data-wrangling, include=FALSE}
# Create a hourly data
# Temperature
temp_hourly <- temp_raw %>%
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  mutate(temp_avg = rowMeans(select(., starts_with("t_ws")), na.rm = TRUE)) %>%
  select(date, hr, temp_avg)
 
head(temp_hourly)

# Aggregate hourly temp to daily mean
temp_daily <- temp_hourly %>%
  mutate(date = as.Date(date)) %>%        
  group_by(date) %>%
  summarise(temp = mean(temp_avg, na.rm = TRUE))

# Humidity
humid_hourly <- humid_raw %>%
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  mutate(humid_avg = rowMeans(select(., starts_with("rh_ws")), na.rm = TRUE)) %>%
  select(date, hr, humid_avg)
 
head(humid_hourly)

# Aggregate hourly humidity to daily mean
humid_daily <- humid_hourly %>%
  mutate(date = as.Date(date)) %>%        
  group_by(date) %>%
  summarise(humidity = mean(humid_avg, na.rm = TRUE))

# Load
load_hourly <- load_raw %>%
  filter(!is.na(meter_id)) %>%          
  mutate(date = as.POSIXct(date, tz = "UTC")) %>%
  select(-meter_id) %>%
  pivot_longer(
    cols      = starts_with("h"),
    names_to  = "hr_str",
    values_to = "load_kwh"
  ) %>%
  mutate(hr = as.integer(sub("h", "", hr_str))) %>%  # "h1" → 1
  select(date, hr, load_kwh) %>%
  arrange(date, hr)
 
head(load_hourly)

# Create an hourly data
hourly_data <- load_hourly %>%
  left_join(temp_hourly, by = c("date", "hr")) %>%
  left_join(humid_hourly,   by = c("date", "hr")) %>%
  arrange(date, hr)

# Create a daily data
daily_data <- hourly_data %>%
  group_by(date) %>%
  summarise(
    load_daily = mean(load_kwh,  na.rm = TRUE),
    temp_avg   = mean(temp_avg, na.rm = TRUE),
    humid_avg  = mean(humid_avg,   na.rm = TRUE),
    .groups    = "drop"
  ) %>%
  arrange(date)

```

```{r preview-daily-data}
knitr::kable(head(daily_data, 10), caption = "First 10 rows of the daily dataset")
```

# 2. Explanatory Data Analysis
An initial time series plot of daily electricity load from January 2005 to June 2010 reveals clear temporal structure characterized by both trend and seasonality. The series exhibits recurring fluctuations, suggesting the presence of strong periodic components, along with an overall upward movement in later years, indicating a gradual increase in load demand over time.

To further decompose these patterns, an STL (Seasonal-Trend decomposition using Loess) was applied with both weekly and annual seasonal periods. The decomposition confirms a pronounced weekly seasonality, likely reflecting differences between weekdays and weekends, as well as a longer annual cycle associated with seasonal demand variations (e.g., weather-driven consumption). The trend component shows a moderate upward trajectory, particularly toward the end of the training period, while the remainder component appears relatively centered around zero, suggesting that most systematic variation has been captured by the trend and seasonal components.

Autocorrelation analysis provides additional insight into the temporal dependence structure. The ACF plot shows a slow decay, indicating strong persistence and non-stationarity in the series, which is consistent with the observed trend and seasonal effects. The PACF plot displays a significant spike at lag 1 followed by smaller contributions at subsequent lags, suggesting that short-term dependencies are important and that lower-order autoregressive components may be sufficient to capture the serial correlation.

```{r explanatory Data Analysis, echo=FALSE}
test_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))

validation_data <- daily_data %>%
  filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

ts_test1 <- msts(test_data$load_daily,
                    seasonal.periods = c(7, 365.25),
                    start = c(2005, 1))

ts_load <- msts(daily_data$load_daily,
                seasonal.periods = c(7, 365.25),
                start = c(2005, 1))

TS_Plot <- 
  ggplot(test_data, aes(x = date, y = load_daily)) +
      geom_line() +
      labs(title = "Daily Electricity Load - Training Data",
           x     = "Date",
           y     = "Load (kWh)") +
      theme_bw()
plot(TS_Plot)

# STL decomposition (more robust than decompose() for msts)
stl_fit <- mstl(ts_test1)
 
cat("\n=== STL Decomposition ===\n")
plot(stl_fit, main = "STL Decomposition — Trend + Weekly + Annual + Remainder")

#ACF and PACF plots
par(mfrow=c(1,2))
ACF_Plot <- Acf(test_data$load_daily, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(test_data$load_daily, lag = 40)
par(mfrow=c(1,1))
```

# 3. Top 5 Models

## (1) Model 1: TBATS
The TBATS model was selected as an initial forecasting approach due to its ability to handle complex seasonal structures, particularly multiple and non-integer seasonal periods. Given that the electricity load data exhibits both weekly and annual  seasonal trends, TBATS provides a flexible framework by incorporating trigonometric representations of seasonality, Box-Cox transformations, and ARMA error structures.

As shown in the results, the TBATS model is able to reasonably track the underlying level and general direction of the series in the validation period. The forecasts reflect the expected seasonal fluctuations and maintain consistency with the recent historical range of the data. However, some deviations between the predicted and actual values are visible, particularly around periods of higher volatility, suggesting that while the model captures broad seasonal patterns effectively, it may be less responsive to short-term irregular spikes.
```{r, echo=FALSE}
## Training and Test
n_valid <- 31  

ts_train <- subset(ts_load, end = length(ts_load) - n_valid)
ts_valid <- subset(ts_load, start = length(ts_load) - n_valid + 1)

length(ts_train)
length(ts_valid) 

## Model 1: TBATS
fit_tbats <- tbats(ts_train)
forecast_tbats <- forecast(fit_tbats, h = n_valid)

accuracy_tbats <- accuracy(forecast_tbats, ts_valid)
print(accuracy_tbats)

# Model 1 Plot
autoplot(forecast_tbats) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 1: TBATS",
       x = "Date", y = "Demand") +
  theme_bw()
```

## (2) Model 2: Fourier + ARIMA


```{r, echo=FALSE}
# Model 2: Fourier + ARIMA
fourier_train <- fourier(ts_train, K = c(2, 4))
fourier_valid <- fourier(ts_train, K = c(2, 4), h = n_valid)

fit_fourier <- auto.arima(ts_train, 
                          xreg = fourier_train,
                          seasonal = FALSE,  
                          stepwise = FALSE,
                          approximation = FALSE)

forecast_fourier <- forecast(fit_fourier, 
                             xreg = fourier_valid,
                             h = n_valid)

accuracy_fourier <- accuracy(forecast_fourier, ts_valid)
print(accuracy_fourier)

# Model 2 Plot
autoplot(forecast_fourier) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 2: Fourier + ARIMA",
       x = "Date", y = "Demand") +
  theme_bw()
```

## (3) Model 3: NNETAR (Neural Network)
The NNETAR (Neural Network Autoregression) model was implemented to capture potential nonlinear patterns in the electricity load series that may not be fully addressed by traditional linear time series models.

As observed in the results, the NNETAR model is able to follow the general level of the series during the validation period; however, its forecasts appear comparatively smoother and less responsive to sharp spikes in demand. While the model captures the underlying structure to some extent, it tends to underrepresent periods of high volatility and peak load.
```{r, echo=FALSE}
# Model 3: NNETAR (Neural Network)
fit_nnetar <- nnetar(ts_train, P = 1, repeats = 20)
forecast_nnetar <- forecast(fit_nnetar, h = n_valid)

accuracy_nnetar <- accuracy(forecast_nnetar, ts_valid)
print(accuracy_nnetar)

# Model 3 Plot
autoplot(forecast_nnetar) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 3: NNETAR",
       x = "Date", y = "Demand") +
  theme_bw()
```

## (4) Model 4: STL + ARIMA
The STL + ARIMA approach combines time series decomposition with a parametric forecasting model, where the series is first decomposed into trend and seasonal components using STL, and an ARIMA model is then fitted to the deseasonalized remainder.

Based on the results, the STL + ARIMA model captures the general level of the series reasonably well; however, its forecasts exhibit substantial variability and wide fluctuations, particularly in the prediction interval. In some instances, the model produces unrealistic dips and spikes, indicating instability in capturing short-term dynamics after decomposition.
```{r, echo=FALSE}
# Model 4: STL + ARIMA
fit_stl_arima <- stlm(ts_train, 
                      s.window = "periodic",
                      method = "arima")

forecast_stl_arima <- forecast(fit_stl_arima, h = n_valid)

accuracy_stl_arima <- accuracy(forecast_stl_arima, ts_valid)
print(accuracy_stl_arima)

# Model 4 Plot
autoplot(forecast_stl_arima) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 4: STL + ARIMA",
       x = "Date", y = "Demand") +
  theme_bw()
```

## (5) Model 5: Fourier+ARIMA + TBATS + NNETAR

```{r, echo=FALSE}
# Model 5: Fourier+ARIMA + TBATS + NNETAR
ensemble_mean <- (forecast_fourier$mean + 
                  forecast_tbats$mean + 
                  forecast_nnetar$mean) / 3

accuracy_ensemble <- accuracy(ensemble_mean, ts_valid)
print(accuracy_ensemble)

# Modeling
fourier_full <- fourier(ts_load, K = c(2, 4))
fourier_future <- fourier(ts_load, K = c(2, 4), h = 31)

fit_fourier_full <- auto.arima(ts_load,
                               xreg = fourier_full,
                               seasonal = FALSE,
                               stepwise = FALSE,
                               approximation = FALSE)

fit_tbats_full <- tbats(ts_load)
fit_nnetar_full <- nnetar(ts_load, P = 1, repeats = 20)

# Forecast
fc_fourier_full <- forecast(fit_fourier_full, xreg = fourier_future, h = 31)
fc_tbats_full   <- forecast(fit_tbats_full, h = 31)
fc_nnetar_full  <- forecast(fit_nnetar_full, h = 31)

# Ensemble Average
ensemble_final <- (fc_fourier_full$mean + 
                   fc_tbats_full$mean + 
                   fc_nnetar_full$mean) / 3
```

## Model Comparison

Through the comparison, Ensemble model has the lowest RMSE. But when submitting to Kaggle, TBATS is the best forecasting model for the competition.

```{r, echo=FALSE}
all_accuracy <- rbind(
  accuracy(ensemble_mean,      ts_valid)[1, ],
  accuracy(forecast_tbats,     ts_valid)[2, ],
  accuracy(forecast_nnetar,    ts_valid)[2, ],
  accuracy(forecast_fourier,   ts_valid)[2, ],
  accuracy(forecast_stl_arima, ts_valid)[2, ]
)

rownames(all_accuracy) <- c("Ensemble (TBATS+Fourier+NNETAR)",
                            "TBATS", 
                            "NNETAR",
                            "Fourier+ARIMA", 
                            "STL+ARIMA")

all_accuracy[ , c("RMSE", "MAPE")] %>%
  knitr::kable(digits = 2, 
               caption = "Model Comparison: Validation Set (July 2010)")
```

# Final

TBATS achieved the best Kaggle public score (0.0998) and was selected as the final model for forecasting daily electricity demand for July 2011.

```{r, echo=FALSE}
# Full Retrain on 2005-01-01 ~ 2011-06-30
fit_tbats_final <- tbats(ts_load)

# July 2011 Forecast (31 days)
fc_tbats_final <- forecast(fit_tbats_final, h = 31)

# Plot
autoplot(fc_tbats_final) +
  labs(title = "Final Forecast: July 2011 (TBATS)",
       x = "Date", y = "Load (kWh)") +
  theme_bw()
```

```{r save-output, include=FALSE}
# CSV
dates_july2011 <- seq(as.Date("2011-07-01"), 
                      as.Date("2011-07-31"), 
                      by = "day")

submission_final <- data.frame(
  date = dates_july2011,
  load = as.numeric(fc_tbats_final$mean)
)

write.csv(submission_final, 
          "Kim_Huang_FinalSubmission.csv", 
          row.names = FALSE)
```

# References: Other Trials
Beyond the five primary models selected for comparison, a range of additional forecasting approaches were explored to better understand the behavior of the series and assess alternative modeling strategies. These included simple baseline methods such as the arithmetic mean and seasonal naïve (SNAIVE) models, as well as naïve and simple exponential smoothing (SES) applied to a deseasonalized series obtained via STL decomposition. These models provided useful reference points and helped evaluate the impact of explicitly removing seasonal components prior to modeling.

More advanced specifications were also considered, including SARIMA and SARIMAX models, with the latter incorporating exogenous variables such as temperature and humidity to account for external drivers of electricity demand. In addition, hybrid approaches such as STL combined with ETS (STL+ETS) and Fourier-based ARIMA models with weather covariates were tested to better capture multiple seasonalities and nonlinear effects. While some of these models demonstrated competitive performance, they were ultimately not selected as primary models due to either lower predictive accuracy on the validation set or reduced stability compared to the final set of chosen models. Please check Rmd file to check all the models that we tried for forecasting.


```{r eval=FALSE, include=FALSE}
decompose_load <- mstl(ts_test1)
plot(decompose_load, main = "STL Decomposition - Daily Load")

# remainder
deseasonal_load <- seasadj(decompose_load)

# Plot deseasonal series
TS_deseas_plot <- 
  ggplot() +
  geom_line(aes(x = test_data$date, y = as.numeric(deseasonal_load))) +
  labs(title = "Deseasonal Daily Load",
       x     = "Date",
       y     = "Load (kWh)") +
  theme_bw()
plot(TS_deseas_plot)

# Model 1: Arithmetic Mean
MEAN_seas <- meanf(y = ts_test1, h = 31)
checkresiduals(MEAN_seas)

# Model 2: Seasonal Naive
SNAIVE_seas <- snaive(ts_test1, h = 31)
checkresiduals(SNAIVE_seas)

# Fitting Models to the Deseasonal Series

# Model 3: Naive on Deseasonal Data
NAIVE_deseas <- naive(deseasonal_load, h = 31)
checkresiduals(NAIVE_deseas)

# Model 4: Simple Exponential Smoothing on Deseasonal Data
SES_deseas <- ses(deseasonal_load, h = 31)
checkresiduals(SES_deseas)

```

```{r eval=FALSE, include=FALSE}
# Comparison with validation data
actual_val <- validation_data$load_daily
rmse <- function(actual, predicted) sqrt(mean((actual - predicted)^2))

rmse_results <- c(
  MEAN        = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE      = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas  = rmse(actual_val, as.numeric(SES_deseas$mean))
)

print(rmse_results)

```

```{r eval=FALSE, include=FALSE}
## Retrain SES_deseas on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start = c(2005, 1))

# Decompose full training series
decompose_full <- mstl(ts_full_training)
deseasonal_full <- seasadj(decompose_full)

# Re-fit SES on deseasonal full training data
SES_final <- ses(deseasonal_full, h = 31)
checkresiduals(SES_final)

# Final forecast for July 1–31, 2011
final_forecast <- data.frame(
  date     = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast = as.numeric(SES_final$mean)
)

print(final_forecast)
```

```{r eval=FALSE, include=FALSE}
## Add Seasonal Component Back to Final Forecast
# Step 1: Extract seasonal components from decomposition
seasonal_weekly <- decompose_full[, "Seasonal7"]
seasonal_annual <- decompose_full[, "Seasonal365.25"]

# Step 2: Find which days in the training data correspond to July
# so we can match the same seasonal pattern for July 2011
full_dates <- seq(as.Date("2005-01-01"), as.Date("2011-06-30"), by = "day")

july_idx <- which(format(full_dates, "%m") == "07" & 
                  format(full_dates, "%d") %in% sprintf("%02d", 1:31))

# Get the LAST July in training data (July 2010) seasonal indices
last_july_idx <- tail(july_idx, 31)

# Step 3: Extract seasonal values for those 31 days
seas_weekly_july <- as.numeric(seasonal_weekly)[last_july_idx]
seas_annual_july <- as.numeric(seasonal_annual)[last_july_idx]

# Step 4: Add back to flat SES forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SES_final$mean),
  seas_weekly    = seas_weekly_july,
  seas_annual    = seas_annual_july,
  forecast       = as.numeric(SES_final$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 5: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SES (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()

# Export for submission
write.csv(final_forecast %>% select(date, forecast),
          "Kim_Huang_FirstSubmission.csv", row.names = FALSE)

# Forecast with SNAIVE on training data (validation check)
SNAIVE_seas <- snaive(ts_test1, h = 31)
```

```{r eval=FALSE, include=FALSE}
## SARIMA & SARIMAX
# Join exogenous variables back to test and validation dataset
test_data <- test_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

validation_data <- validation_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

# Build xreg matrix
xreg_train <- cbind(
  temp     = test_data$temp,
  humidity = test_data$humidity
)

xreg_val <- cbind(
  temp     = validation_data$temp,
  humidity = validation_data$humidity
)
```

```{r eval=FALSE, include=FALSE}
## SARIMA on Deseasonal Series
SARIMA_model <- auto.arima(
  deseasonal_load,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_model)

SARIMA_fc <- forecast(SARIMA_model, h = 31)

# Model 6: SARIMAX on Deseasonal Series
SARIMAX_model <- auto.arima(
  deseasonal_load,
  xreg          = xreg_train,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMAX_model)

SARIMAX_fc <- forecast(SARIMAX_model, h = 31, xreg = xreg_val)
```

```{r eval=FALSE, include=FALSE}
# Model comparison
# Updated RMSE Comparison
rmse_results <- c(
  MEAN         = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE       = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas   = rmse(actual_val, as.numeric(SES_deseas$mean)),
  SARIMA       = rmse(actual_val, as.numeric(SARIMA_fc$mean)),
  SARIMAX      = rmse(actual_val, as.numeric(SARIMAX_fc$mean))
)

print(sort(rmse_results))
```

```{r eval=FALSE, include=FALSE}
# Apparently, SARIMA model has the lowest RMSE, so we will forecast with SARIMA model.
## Retrain SARIMA on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start            = c(2005, 1))

# Decompose full training series
decompose_full    <- mstl(ts_full_training)
deseasonal_full   <- seasadj(decompose_full)

# Re-fit SARIMA on deseasonal full training data
SARIMA_final <- auto.arima(
  deseasonal_full,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_final)

# Final deseasonal forecast for July 1-31, 2011
SARIMA_final_fc <- forecast(SARIMA_final, h = 31)
```

```{r eval=FALSE, include=FALSE}
# Step 5: Add seasonal components back to SARIMA forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SARIMA_final_fc$mean),
  seas_weekly     = seas_weekly_july,
  seas_annual     = seas_annual_july,
  forecast        = as.numeric(SARIMA_final_fc$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 6: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SARIMA (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x     = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()
```

```{r eval=FALSE, include=FALSE}
# Re-train on FULL data (2005/01/01 ~ 2011/06/30) for final submission
full_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full <- msts(full_data$load_daily,
                seasonal.periods = c(7, 365.25),
                start = c(2005, 1))

SNAIVE_full <- snaive(ts_full, h = 31)
```

```{r eval=FALSE, include=FALSE}
## STL + ETS
fit_stl_ets <- stlf(ts_train, h = n_valid)
forecast_stl_ets <- forecast(fit_stl_ets, h = n_valid)

accuracy_stl_ets <- accuracy(forecast_stl_ets, ts_valid)
print(accuracy_stl_ets)

#Plot
autoplot(forecast_stl_ets) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model: STL + ETS",
       x = "Date", y = "Demand") +
  theme_bw()


## Fourier + ARIMA + Temp/Humidity
# Add Temp and Humidity Data
xreg_train <- daily_data %>%
  slice(1:(nrow(daily_data) - n_valid)) %>%
  select(daily_temp, daily_humid) %>%
  as.matrix()

xreg_valid <- daily_data %>%
  slice((nrow(daily_data) - n_valid + 1):nrow(daily_data)) %>%
  select(daily_temp, daily_humid) %>%
  as.matrix()

nrow(xreg_train)
nrow(xreg_valid)

# Modeling: Fourier + ARIMA + Temp/Humidity
xreg_train_full <- cbind(fourier(ts_train, K = c(2, 4)), xreg_train)
xreg_valid_full <- cbind(fourier(ts_train, K = c(2, 4), h = n_valid), xreg_valid)

fit_fourier_weather <- auto.arima(ts_train,
                                  xreg = xreg_train_full,
                                  seasonal = FALSE,
                                  stepwise = FALSE,
                                  approximation = FALSE)

forecast_fourier_weather <- forecast(fit_fourier_weather,
                                     xreg = xreg_valid_full,
                                     h = n_valid)

accuracy_weather <- accuracy(forecast_fourier_weather, ts_valid)
print(accuracy_weather)

# Plot: Fourier + ARIMA + Temp/Humidity
autoplot(forecast_fourier_weather) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Fourier + ARIMA + Temp/Humid",
       x = "Date", y = "Demand") +
  theme_bw()

## Weighted Average of Tbats and Fourier --> Turned out this is not the best
weights3 <- list(
  c(0.2, 0.8),
  c(0.1, 0.9),
  c(0.0, 1.0)
)

for (w in weights3) {
  ens <- w[1] * forecast_tbats$mean + w[2] * forecast_fourier$mean
  mape <- accuracy(ens, ts_valid)["Test set", "MAPE"]
  cat("TBATS:", w[1], "| Fourier:", w[2], "| MAPE:", round(mape, 4), "\n")
}
```

AI statement: AI tool was used to help us change the format of date in our data wrangling process.
