Forecasting Competition 
Spring 2026
Author: Bella Huang & Yeeun Kim

GitHub repository: https://github.com/BellaHuang03/KimHuang_ENV797_TSA_ForecastCompetition_S26

1. Set up
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(forecast)
library(smooth)
library(ggplot2)
library(tidyverse)
```

2. Import datasets
```{r}
#Load Data
temp_raw <- read_excel("temperature.xlsx")
humid_raw   <- read_excel("relative_humidity.xlsx")
load_raw <- read_excel("load.xlsx")
```

3. Data Wrangling
```{r}
# Load Daily
load_daily <- load_raw %>%
  select(-meter_id) %>%                    
  rowwise() %>%
  mutate(daily_load = mean(c_across(h1:h24), na.rm = TRUE)) %>%
  ungroup() %>%
  select(date, daily_load) %>%
  mutate(date = as.Date(date))

# Temperature Daily
temp_daily <- temp_raw %>%
  mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(daily_temp = mean(c_across(t_ws1:t_ws28), na.rm = TRUE))

# Humidity Daily
humid_daily <- humid_raw %>%
  mutate(date = as.Date(date)) %>%
  group_by(date) %>%
  summarise(daily_humid = mean(c_across(rh_ws1:rh_ws28), na.rm = TRUE))

# Daily Data
daily_data <- load_daily %>%
  left_join(temp_daily, by = "date") %>%
  left_join(humid_daily, by = "date")

head(daily_data)
nrow(daily_data) 
```

4. Exploratory Data Analysis
```{r}
# msts
ts_load <- msts(daily_data$daily_load, 
                seasonal.periods = c(7, 365.25),
                start = c(2005, 1))

# plots
autoplot(ts_load) +
  labs(title = "Daily Electricity Demand (2005–2011)",
       x = "Date", y = "Load (kWh)") +
  theme_bw()


# ACF & PACF
par(mfrow=c(1,2))
Acf(daily_data$daily_load, lag.max = 60, main = "ACF - Daily Demand")
Pacf(daily_data$daily_load, lag.max = 60, main = "PACF - Daily Demand")
```

5. Training and Test
```{r}
n_valid <- 31  

ts_train <- subset(ts_load, end = length(ts_load) - n_valid)
ts_valid <- subset(ts_load, start = length(ts_load) - n_valid + 1)

length(ts_train)
length(ts_valid) 

# Model 1: STL + ETS
fit_stl_ets <- stlf(ts_train, h = n_valid)
forecast_stl_ets <- forecast(fit_stl_ets, h = n_valid)

accuracy_stl_ets <- accuracy(forecast_stl_ets, ts_valid)
print(accuracy_stl_ets)

# Model 1 Plot
autoplot(forecast_stl_ets) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 1: STL + ETS",
       x = "Date", y = "Demand") +
  theme_bw()

# Model 2: TBATS
fit_tbats <- tbats(ts_train)
forecast_tbats <- forecast(fit_tbats, h = n_valid)

accuracy_tbats <- accuracy(forecast_tbats, ts_valid)
print(accuracy_tbats)

# Model 2 Plot
autoplot(forecast_tbats) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 2: TBATS",
       x = "Date", y = "Demand") +
  theme_bw()

# Model 3: Fourier + ARIMA
fourier_train <- fourier(ts_train, K = c(2, 4))
fourier_valid <- fourier(ts_train, K = c(2, 4), h = n_valid)

fit_fourier <- auto.arima(ts_train, 
                          xreg = fourier_train,
                          seasonal = FALSE,  
                          stepwise = FALSE,
                          approximation = FALSE)

forecast_fourier <- forecast(fit_fourier, 
                             xreg = fourier_valid,
                             h = n_valid)

accuracy_fourier <- accuracy(forecast_fourier, ts_valid)
print(accuracy_fourier)

# Model 3 Plot
autoplot(forecast_fourier) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 3: Fourier + ARIMA",
       x = "Date", y = "Demand") +
  theme_bw()

# Model 4: NNETAR (Neural Network)
fit_nnetar <- nnetar(ts_train, P = 1, repeats = 20)
forecast_nnetar <- forecast(fit_nnetar, h = n_valid)

accuracy_nnetar <- accuracy(forecast_nnetar, ts_valid)
print(accuracy_nnetar)

# Model 4 Plot
autoplot(forecast_nnetar) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 4: NNETAR",
       x = "Date", y = "Demand") +
  theme_bw()

# Model 5: STL + ARIMA
fit_stl_arima <- stlm(ts_train, 
                      s.window = "periodic",
                      method = "arima")

forecast_stl_arima <- forecast(fit_stl_arima, h = n_valid)

accuracy_stl_arima <- accuracy(forecast_stl_arima, ts_valid)
print(accuracy_stl_arima)

# Model 5 Plot
autoplot(forecast_stl_arima) +
  autolayer(ts_valid, series = "Actual") +
  labs(title = "Model 5: STL + ARIMA",
       x = "Date", y = "Demand") +
  theme_bw()
```

6. Forecast
```{r}
# ══════════════════════════════════════════
# 6단계: Full Retrain (2005-01-01 ~ 2011-06-30)
# ══════════════════════════════════════════

# 전체 데이터로 msts 재생성 (ts_load 그대로 사용)

# Best Model: Fourier + ARIMA 
fourier_full <- fourier(ts_load, K = c(2, 4))
fourier_future <- fourier(ts_load, K = c(2, 4), h = 31)

fit_fourier_full <- auto.arima(ts_load,
                               xreg = fourier_full,
                               seasonal = FALSE,
                               stepwise = FALSE,
                               approximation = FALSE)

# Forecast: July 2011 
forecast_final <- forecast(fit_fourier_full,
                           xreg = fourier_future,
                           h = 31)

# Plot
autoplot(forecast_final) +
  labs(title = "Final Forecast: July 2011",
       x = "Date", y = "Demand") +
  theme_bw()

# ══════════════════════════════════════════
# 7단계: CSV 제출 파일 생성
# ══════════════════════════════════════════

# July 2011 날짜 생성
dates_july2011 <- seq(as.Date("2011-07-01"), 
                      as.Date("2011-07-31"), 
                      by = "day")

# 제출 파일 생성
submission <- data.frame(
  date = dates_july2011,
  load = as.numeric(forecast_final$mean)
)

# 확인
print(submission)

# CSV 저장
write.csv(submission, 
          "submission_fourier_arima.csv", 
          row.names = FALSE)
```


4. Initial plot - filtering for Jan 2005 to June 2010
```{r}
test_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))

validation_data <- daily_data %>%
  filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

ts_test1 <- msts(test_data$load_daily,
                    seasonal.periods = c(7, 365.25),
                    start = c(2005, 1))
TS_Plot <- 
  ggplot(test_data, aes(x = date, y = load_daily)) +
      geom_line() +
      labs(title = "Daily Electricity Load - Training Data",
           x     = "Date",
           y     = "Load (kWh)") +
      theme_bw()
plot(TS_Plot)

# STL decomposition (more robust than decompose() for msts)
stl_fit <- mstl(ts_test1)
 
cat("\n=== STL Decomposition ===\n")
plot(stl_fit, main = "STL Decomposition — Trend + Weekly + Annual + Remainder")

#ACF and PACF plots
par(mfrow=c(1,2))
ACF_Plot <- Acf(test_data$load_daily, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(test_data$load_daily, lag = 40)
par(mfrow=c(1,1))
```

5. Decompose, forecast with seasonal Naive model and averaging
```{r}
decompose_load <- mstl(ts_test1)
plot(decompose_load, main = "STL Decomposition - Daily Load")

# remainder
deseasonal_load <- seasadj(decompose_load)

# Plot deseasonal series
TS_deseas_plot <- 
  ggplot() +
  geom_line(aes(x = test_data$date, y = as.numeric(deseasonal_load))) +
  labs(title = "Deseasonal Daily Load",
       x     = "Date",
       y     = "Load (kWh)") +
  theme_bw()
plot(TS_deseas_plot)
```
```{r}
# Model 1: Arithmetic Mean
MEAN_seas <- meanf(y = ts_test1, h = 31)
checkresiduals(MEAN_seas)

# Model 2: Seasonal Naive
SNAIVE_seas <- snaive(ts_test1, h = 31)
checkresiduals(SNAIVE_seas)

# Fitting Models to the Deseasonal Series

# Model 3: Naive on Deseasonal Data
NAIVE_deseas <- naive(deseasonal_load, h = 31)
checkresiduals(NAIVE_deseas)

# Model 4: Simple Exponential Smoothing on Deseasonal Data
SES_deseas <- ses(deseasonal_load, h = 31)
checkresiduals(SES_deseas)

```
Comparison with validation data
```{r}
actual_val <- validation_data$load_daily
rmse <- function(actual, predicted) sqrt(mean((actual - predicted)^2))

rmse_results <- c(
  MEAN        = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE      = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas  = rmse(actual_val, as.numeric(SES_deseas$mean))
)

print(rmse_results)

```

```{r}
## Retrain SES_deseas on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start = c(2005, 1))

# Decompose full training series
decompose_full <- mstl(ts_full_training)
deseasonal_full <- seasadj(decompose_full)

# Re-fit SES on deseasonal full training data
SES_final <- ses(deseasonal_full, h = 31)
checkresiduals(SES_final)

# Final forecast for July 1–31, 2011
final_forecast <- data.frame(
  date     = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast = as.numeric(SES_final$mean)
)

print(final_forecast)
```
```{r}
## Add Seasonal Component Back to Final Forecast
# Step 1: Extract seasonal components from decomposition
seasonal_weekly <- decompose_full[, "Seasonal7"]
seasonal_annual <- decompose_full[, "Seasonal365.25"]

# Step 2: Find which days in the training data correspond to July
# so we can match the same seasonal pattern for July 2011
full_dates <- seq(as.Date("2005-01-01"), as.Date("2011-06-30"), by = "day")

july_idx <- which(format(full_dates, "%m") == "07" & 
                  format(full_dates, "%d") %in% sprintf("%02d", 1:31))

# Get the LAST July in training data (July 2010) seasonal indices
last_july_idx <- tail(july_idx, 31)

# Step 3: Extract seasonal values for those 31 days
seas_weekly_july <- as.numeric(seasonal_weekly)[last_july_idx]
seas_annual_july <- as.numeric(seasonal_annual)[last_july_idx]

# Step 4: Add back to flat SES forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SES_final$mean),
  seas_weekly    = seas_weekly_july,
  seas_annual    = seas_annual_july,
  forecast       = as.numeric(SES_final$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 5: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SES (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()

# Export for submission
write.csv(final_forecast %>% select(date, forecast),
          "Kim_Huang_FirstSubmission.csv", row.names = FALSE)

```
<<<<<<< HEAD

Another try
```{r}
# Forecast with SNAIVE on training data (validation check)
SNAIVE_seas <- snaive(ts_test1, h = 31)
=======
More models to try: SARIMA & SARIMAX
```{r}
# Join exogenous variables back to test and validation dataset
test_data <- test_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

validation_data <- validation_data %>%
  left_join(temp_daily,  by = "date") %>%
  left_join(humid_daily, by = "date")

# Build xreg matrix
xreg_train <- cbind(
  temp     = test_data$temp,
  humidity = test_data$humidity
)

xreg_val <- cbind(
  temp     = validation_data$temp,
  humidity = validation_data$humidity
)
```

```{r}
# Model 5: SARIMA on Deseasonal Series
SARIMA_model <- auto.arima(
  deseasonal_load,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_model)

SARIMA_fc <- forecast(SARIMA_model, h = 31)

# Model 6: SARIMAX on Deseasonal Series
SARIMAX_model <- auto.arima(
  deseasonal_load,
  xreg          = xreg_train,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMAX_model)

SARIMAX_fc <- forecast(SARIMAX_model, h = 31, xreg = xreg_val)
```
Model comparison
```{r}
# Updated RMSE Comparison
rmse_results <- c(
  MEAN         = rmse(actual_val, as.numeric(MEAN_seas$mean)),
  SNAIVE       = rmse(actual_val, as.numeric(SNAIVE_seas$mean)),
  NAIVE_deseas = rmse(actual_val, as.numeric(NAIVE_deseas$mean)),
  SES_deseas   = rmse(actual_val, as.numeric(SES_deseas$mean)),
  SARIMA       = rmse(actual_val, as.numeric(SARIMA_fc$mean)),
  SARIMAX      = rmse(actual_val, as.numeric(SARIMAX_fc$mean))
)

print(sort(rmse_results))
```

Apparently, SARIMA model has the lowest RMSE, so we will forecast with SARIMA model.

```{r}
## Retrain SARIMA on Full Dataset
full_training_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full_training <- msts(full_training_data$load_daily,
                         seasonal.periods = c(7, 365.25),
                         start            = c(2005, 1))

# Decompose full training series
decompose_full    <- mstl(ts_full_training)
deseasonal_full   <- seasadj(decompose_full)

# Re-fit SARIMA on deseasonal full training data
SARIMA_final <- auto.arima(
  deseasonal_full,
  seasonal      = FALSE,
  stepwise      = FALSE,
  approximation = FALSE
)
checkresiduals(SARIMA_final)

# Final deseasonal forecast for July 1-31, 2011
SARIMA_final_fc <- forecast(SARIMA_final, h = 31)
```

```{r}
# Step 5: Add seasonal components back to SARIMA forecast
final_forecast <- data.frame(
  date           = seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day"),
  forecast_deseas = as.numeric(SARIMA_final_fc$mean),
  seas_weekly     = seas_weekly_july,
  seas_annual     = seas_annual_july,
  forecast        = as.numeric(SARIMA_final_fc$mean) + seas_weekly_july + seas_annual_july
)

print(final_forecast)

# Step 6: Plot to verify variation is restored
ggplot(final_forecast, aes(x = date)) +
  geom_line(aes(y = forecast_deseas, color = "Flat SARIMA (deseasonal)")) +
  geom_line(aes(y = forecast,        color = "Final (with seasonality)")) +
  labs(title = "July 2011 Final Forecast",
       x     = "Date", y = "Load (kWh)", color = "Series") +
  theme_bw()

# Export for submission
write.csv(final_forecast %>% select(date, forecast),
          "Kim_Huang_SecondSubmission.csv", row.names = FALSE)
```

AI statement: AI tool was used to help us change the format of date in our data wrangling process.

>>>>>>> 08bd667d68fa7ce15df76827ec7ce607ed6edbc7

# Re-train on FULL data (2005/01/01 ~ 2011/06/30) for final submission
full_data <- daily_data %>%
  filter(date >= as.Date("2005-01-01") & date <= as.Date("2011-06-30"))

ts_full <- msts(full_data$load_daily,
                seasonal.periods = c(7, 365.25),
                start = c(2005, 1))

SNAIVE_full <- snaive(ts_full, h = 31)

# Build submission file
july_2011 <- seq(as.Date("2011-07-01"), as.Date("2011-07-31"), by = "day")

submission <- data.frame(
  date = as.Date(july_2011),
  load = round(as.numeric(SNAIVE_full$mean), 2)
)

write.csv(submission, "submission_second.csv", row.names = FALSE)
print(submission)
```